\section{The Born-Oppenheimer Approximation}

We begin by writing the Hamiltonian for this system; it's just the combination of the kinetic and potential energies:

\begin{equation}
  \hat{H} = \frac{\hat{p}_1^2}{2M} + \frac{\hat{p}_2^2}{2m} + \frac{1}{2}k_1\hat{x}_1^2 + \frac{1}{2}k_2\hat{x}_2^2 + \frac{1}{2}k_{12}(\hat{x}_1 - \hat{x}_2)^2.
\end{equation}

To start with the Born-Oppenheimer approximation, we first make the assumption that the heavier particle is so much heavier that we can neglect its motion compared to the lighter particle. In this case, this means that its kinetic energy is significantly smaller than that of the lighter particle. We also recognize that the Hamiltonian can be split up into three parts: $\hat{H} = \hat{H}_1 + \hat{H}_{12} + \hat{H}_{2}$ where $\hat{H}_i$ is the Hamiltonian of the $i$th particle and $\hat{H}_{12}$ is the interaction part containing the last potential term. With our assumption, we have that $\hat{H}_1 << \hat{H}_{12} + \hat{H}_2$, so we solve the latter:

\begin{equation}
  (\hat{H}_{12} + \hat{H}_2)\ket{\psi_n} = (E_2)_n\ket{\psi_n},
\end{equation}

where $\ket{\psi_n}$ is the $n$th eigenket for the light particle. This is something that we can solve. Writing out this new Hamiltonian, we have

\begin{equation}
  \hat{H}_{12} + \hat{H}_2 = \frac{\hat{p}_2^2}{2m} + \frac{1}{2}k_2x_2^2 + \frac{1}{2}k_{12}(\hat{x}_1 - \hat{x}_2)^2.
\end{equation}

We can complete the square here and get a decently simple equation:

\begin{align}
  \hat{H}_{12} + \hat{H}_2 &= \frac{\hat{p}_2^2}{2m} + \frac{1}{2}k_2x_2^2 + \frac{1}{2}k_{12}x_1^2 - k_{12}x_1x_2 + \frac{1}{2}k_{12}x_2^2 \\
                           &= \frac{\hat{p}_2^2}{2m} + \frac{1}{2}(k_2 + k_{12})x_2^2 - k_{12}x_1x_2 + \frac{1}{2}k_{12}x_1^2 \\
                           &= \frac{\hat{p}_2^2}{2m} + \frac{1}{2}(k_2 + k_{12})\left[ x_2^2 - \left( \frac{2k_{12}}{k_2 + k_{12} x_1}x_2 + \frac{k_{12}}{k_2 + k_12}x_1^2 \right) \right].
\end{align}

Defining $a \equiv k_{12}/(k_2 + k_{12})$ to make things simpler, we have

\begin{align}
  \hat{H}_{12} + \hat{H}_2 &= \frac{\hat{p}_2^2}{2m} + \frac{1}{2}(k_2 + k_{12})\left[ x_2^2 - 2ax_1x_2 + ax_1^2 \right] \\
                           &= \frac{\hat{p}_2^2}{2m} + \frac{1}{2}(k_2 + k_{12})\left[ (x_2 - ax_1)^2 + a(1-a)x_1^2 \right] \\
                           &= \frac{\hat{p}_2^2}{2m} + \frac{1}{2}(k_2 + k_{12})(x_2 - ax_1)^2 + k_{12}(1-a)x_1^2.
\end{align}

We have recovered a normal simple harmonic oscillator Hamiltonian, but with $k'_2 = k_2 + k_{12}$ and a constant term. The constant term won't do anything but sum the normal simple harmonic oscillator solutions. So:

\begin{equation}
  (E_2)_0 = \hbar\omega'_2\left( n + \frac{1}{2} \right) + k_{12}(1-a)x_1^2 = \hbar\omega'_2\left( n + \frac{1}{2} \right) + x_1^2\left( \frac{k_2k_{12}}{k_2 + k_{12}} \right),
\end{equation}

where $\omega'_2 = \sqrt{k'_2/m} = \sqrt{(k_2 + k_{12})/m}$.

We now adopt the product ans\"atz, that is, that we express the total wavefunction as a product of the heavier and lighter particle wavefunctions:

\begin{equation}
  \ket{\Psi} = \ket{\psi}\ket{\phi}, \quad\text{or}\quad \Psi(x_1,x_2) = \psi(x_1)\phi(x_2).
\end{equation}

To look at the wavefunction for the heavier particle, we make another assumption: the lighter particle remains in the same state as the system progresses. If this is the case, we can just place it in the ground state without loss of generality. Now, we act on the total wavefunction with $\hat{H}_{12} + \hat{H}_2$. It will ignore the heavy part since this combined light operator sees it as essentially constant. This leaves us with:

\begin{equation}
  (\hat{H}_{12} + \hat{H}_2)\ket{\Psi} = (E_2)_0\ket{\psi}\ket{\phi}.
\end{equation}

Now if we act with $\hat{H}_1$ on this Hamiltonian, we can first consider the kinetic term:

\begin{align}
  \hat{p}_1^2\ket{\psi}\ket{\phi} &= (-\hbar^2\partial_{x_1}^2 \ket{\psi})\ket{\phi} + \ket{\psi}(\hat{p}_1^2 \ket{\phi}) \\ 
                                          & -2i\hbar (\partial_{x_1} \ket{\psi})(\hat{p}_1\ket{\phi})
\end{align}

If we just define $U_1(x_1) \equiv kx_1^2/2$, then the entire $\hat{H}_1$ term acting on the total wavefunction looks like

\begin{align}
  \hat{H}_1\ket{\Psi} &= \left( -\frac{\hbar^2}{2M}\partial_{x_1}^2 \ket{\psi} \right)\ket{\phi} + \ket{\psi}\left( \frac{\hat{p}_1^2}{2M} \ket{\phi} \right) \\
                      & - \frac{i\hbar}{m}(\partial_{x_1}\ket{\psi})(\hat{p}_1 \ket{\phi}) + \ket{\psi} U(x_1) \ket{\phi}.
\end{align}

So, adding this with the light-particle solution (which is just the energy eigenvalues) we get the action of the total Hamiltonian on the total wavefunction:

\begin{align}                        
  \hat{H}\ket{\Psi} &= \ket{\psi}\left[ - \frac{\hbar^2}{2M}\partial_{x_1}^2 + U_1(x_1) + (E_2)_0 \right] \ket{\phi} \\
                    & - \left[ \frac{i\hbar}{M}(\partial_{x_1} \ket{\psi})\hat{p}_1 + \frac{\hbar^2}{2M} \partial_{x_1}^2 \ket{\psi} \right]\ket{\phi}.
\end{align}

We can now integrate out the light-particle degrees of freedom by acting on the left with $\bra{\psi}$:

\begin{align}
  \braket{\psi | \hat{H} | \Psi} &= \bra{\psi} \left[ - \frac{\hbar^2}{2M}\partial_{x_1}^2 + U_1(x_1) + (E_2)_0 \right]\ket{\phi} \\
                                       & - \left[ \frac{i\hbar}{m}(\braket{\psi | \partial_{x_1} | \psi}\hat{p}_1 + \frac{\hbar^2}{2M}(\braket{\psi | \partial_{x_1}^2 | \psi}) \right] \ket{\phi}.
\end{align}

We know from class that the two terms in the second line are zero; the first is zero due to time-reversal symmetry, the second is proportional to the kinetic energy of the lighter particle divided by the mass of the large particle, which in our approximation, we can take to zero. Thus,

\begin{equation}
  \braket{\psi | \hat{H} | \Psi} = \bra{\psi}\left[ \frac{\hat{p}_1^2}{2M} + U_1(x_1) + (E_2)_0 \right]\ket{\phi},
\end{equation}

hence, we can define an effective Hamiltonian for the heavy particle:

\begin{equation}
  \hat{H}_{1,\mathrm{eff}} = \frac{\hat{p}_1^2}{2M} + U_1(x_1) + (E_2)_0 = \frac{\hat{p}_1^2}{2M} + \frac{1}{2}k_1x_1^2 + (E_2)_0.
\end{equation}

This is an easily Hamiltonian to solve. First, the ground state of the lighter particle has

\begin{equation}
  (E_2)_0 = \frac{\hbar\omega_2'}{2} + x_1^2\left( \frac{k_2k_{12}}{k_2 + k_{12}} \right),
\end{equation}

which leads to

\begin{equation}
  \hat{H}_{1,\mathrm{eff}} = \frac{\hat{p}_1^2}{2M} + \frac{1}{2}\left( k_1 + \frac{2k_2k_{12}}{k_2 + k_{12}} \right)x_1^2 + \frac{\hbar\omega_2'}{2}.
\end{equation}

This is just another SHO with a constant term. So, defining $k_1'$ as the stuff inside the parentheses, we have

\begin{equation}
  (E_1)_n = \hbar\omega_1'\left( n + \frac{1}{2} \right) + \frac{\hbar\omega_2'}{2},
\end{equation}

where $\omega_1' = \sqrt{k_1'/M}$.

At this point, we have solved this problem in the Born-Oppenheimer approximation. The next order of business would be to solve it exactly, i.e. without any approximations. This is something I still don't quite understand, so I will just recap my very basic understanding. First, we can better express this in matrix form like so:

\begin{equation}
  H = \frac{1}{2}\mathbf{P}^\intercal \mathbf{M}^{-1} \mathbf{P} + \frac{1}{2}\mathbf{X}^\intercal \mathbf{K} \mathbf{X},
\end{equation}

where $\mathbf{P} = \begin{pmatrix}\hat{p}_1 & \hat{p}_2\end{pmatrix}$, $\mathbf{M}$ is a positive-definite, diagonal mass matrix, $\mathbf{X} = \begin{pmatrix}x_1 & x_2\end{pmatrix}$, and

\begin{equation}
  \mathbf{K} = \begin{pmatrix}k_1 + k_{12} & -k_{12} \\ -k_{12} & k_1 + k_{12}\end{pmatrix}.
\end{equation}

(The next half-page or so is overexplained, but I am doing it in hopes that by explaining it thoroughly on paper, it will become more clear to me. It will also make it easier for you to point out where I inevitably went wrong somewhere.) The idea is to diagonalize $\mathbf{K}$, since that way we would have two decoupled harmonic oscillators, whose energies we can very easily solve for (it'd just be a sum of the two). However, we can't just diagonalize $\mathbf{K}$, since we are in essence doing a coordinate transformation, meaning the momentum matrices would have to be transformed non-trivially as well. In doing so, in order for the kinetic part of the Hamiltonian to remain diagonal, the coefficients must have mass-dependence.

This can be seen by considering the kinetic part only. If we were to do a coordinate transformation, we end up with:

\begin{equation}
  \frac{1}{2}\mat{P}^\intercal \mat{M}^{-1} \mat{P} \quad\rightarrow\quad \frac{1}{2}\mat{P}^\intercal \mat{U}^{-1} \mat{M}^{-1} \mat{U} \mat{P}.
\end{equation}

Expressing elements of $\mat{U}$ generally as $u_{ij}$ and elements of $\mat{U}_{-1}$ generally as $u^{-1}_{ij}$, we find that an off-diagonal component of the final matrix sandwiched between the momentum vectors would look like:

\begin{equation}
  \left(\mat{U}^{-1}\mat{M}^{-1}\mat{U}\right)_{12} = \frac{u^{-1}_{11}u_{12}}{m_1} + \frac{u^{-1}_{12}u_{22}}{m_2}.
\end{equation}

Without any mass dependence, there is no way for these terms to cancel. Hence, we can consider splitting the mass matrix up into $\mat{U}^{-1}\mat{M}^{-1/2}\mat{M}^{-1/2}\mat{U}$, such that $\mat{U}^{-1}\mat{M}^{-1/2}$ and $\mat{M}^{-1/2}\mat{U}$ are the two transformations, and in this way, the kinetic term remains trivially diagonal and we can look to the potential term to make any determinations. Under these transformations, the potential term will look $\mat{x}^\intercal\mat{U}^{-1}\mat{M}^{-1/2}\mat{K}\mat{M}^{-1/2}\mat{U}\mat{x}$. At the end of the day, then, we are looking to diagonalize not $\mathbf{K}$ but $\mathbf{M}^{-1/2}\mathbf{K}\mathbf{M}^{-1/2}$:

\begin{equation}
  \mathbf{M}^{-1/2}\mathbf{K}\mathbf{M}^{-1/2} = \begin{pmatrix}\frac{k_1 + k_{12}}{M} & - \frac{k_{12}}{\sqrt{Mm}} \\ - \frac{k_{12}}{\sqrt{Mm}} & \frac{k_1 + k_{12}}{m}\end{pmatrix}.
\end{equation}

I don't want to type everything out here, but by diagonalizing this and using the ever-wonderful quadratic formula, we arrive the eigenvalues, which are the two frequencies (squared) of the normal modes:

\begin{equation}
  \omega^2_\pm = \frac{-\frac{k_{12}(M + m) + mk_1 + Mk_2}{Mm} \pm \sqrt{\left( \frac{k_{12}(M + m) + mk_1 + Mk_2}{Mm} \right)^2 - 4\left( \frac{k_1k_{12} + k_1k_2 + k_{12}k_2}{Mm} \right)}}{2}.
\end{equation}

Since we have $m << M$, we can simplify the term outside the square root (and the identical term that's squared inside the square root and get

\begin{equation}
  \omega^2_\pm = \frac{- \frac{k_{12} + k_2}{m} \pm \sqrt{\left( \frac{k_{12} + k_2}{m} \right)^2 - 4\left( \frac{k_1k_{12} + k_1k_2 + k_{12}k_2}{Mm} \right)}}{2}.
\end{equation}

Further, if we multiply top and bottom by $m$, we get

\begin{equation}
  \omega^2_\pm = \frac{-(k_{12} + k2) \pm \sqrt{(k_{12} + k_2)^2 - 4 \frac{m}{M}(k_1k_{12} + k_1k_2 + k_{12}k_2)}}{2m},
\end{equation}

and the second term in brackets vanishes. If we then take the limiting case $k_{12} << k_1,k_2$, the $\pm$ gives two terms, one of which is zero, and all we get left is $\omega = \sqrt{k_2/m}$ (taking the positive root), which gives us energy levels which makes complete sense! If one particle is super heavy and doesn't move much compared to the other, and they hardly interact, then we essentially just have the one oscillator, which is exactly what we just got.

We also see pretty easily that the same occurs for our approximation case.




%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: "../../HW2"
%%% End:
